---
title: "IAV Key Exposures"
author: "jonasalexandersteiner"
date: "`r format(Sys.Date())`"
output: 
  html_document:
    toc: false
    number_sections: false
    code_folding: hide
    df_print: paged
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
library(pROC)
library(car)
library(kableExtra)
output_dir <- "C:/Users/js22i117/OneDrive - Universitaet Bern/Dokumente/SIV_Projekt/Data_processing/SIV_farms100/04_output"
load(file.path(output_dir, "SIV_logistic_regression_results.RData"))
# Helper: HTML decode function for variable names
decode_html <- function(x) {
  x <- gsub("&lt;", "<", x, fixed = TRUE)
  x <- gsub("&gt;", ">", x, fixed = TRUE)
  x <- gsub("&ge;", "≥", x, fixed = TRUE)
  x <- gsub("&le;", "≤", x, fixed = TRUE)
  x
}
```

# Logistic regression of key factors in binary PLSR

This report presents summary results and visualizations for a logistic regression model for IAV status in pig herds:

- Top 25 predictors by PLS-DA VIP were identified (husbandry subset).
- Univariate logistic regressions were computed for these predictors to assess their invidual influence on IAV detection.
- The final 4 predictors were chosen based on greatest VIP scores, then manually screened for biological plausibility. Variables reflecting intensive farming, missingness, or non-independence/confounding were excluded (see column 'exclusion_due_to_biological_plausibility' in tables).  
- Multivariable logistic regression was performed with diagnostic checks and transparent reporting.
- Calibration of the model was assessed using the Hosmer-Lemeshow test and a calibration plot.

---

## Husbandry Subset Model

### Model Sample Size

```{r results='asis'}
cat("N (Husbandry model used):", nrow(df_logit_husbandry), "<br>")
```

### Top 25 Predictors: Univariate OR Table

**Top 25 predictors by PLS-DA VIP, with univariate odds ratios, 95% confidence intervals, VIP scores, and biological exclusion rationale. Effect size is VIP (distance from null effect). Variables excluded for biological reasons are annotated.**

```{r results='asis'}
if (exists("vip_table_husbandry")) {
  vip_table_husbandry_disp <- as.data.frame(vip_table_husbandry)
  vip_table_husbandry_disp <- vip_table_husbandry_disp[, !grepl("^x(\\.{3}|…)[0-9]+$", names(vip_table_husbandry_disp))]
  vip_table_husbandry_disp <- vip_table_husbandry_disp[, c("variable", "VIP", "exclusion_due_to_biological_plausibility")]
  vip_table_husbandry_disp$variable <- decode_html(vip_table_husbandry_disp$variable)
  vip_table_husbandry_disp <- vip_table_husbandry_disp %>%
    mutate(VIP = signif(VIP, 3)) %>%
    arrange(desc(VIP))
  kable(vip_table_husbandry_disp, digits=3, caption="Top 25 by VIP with exclusion rationale", align = "l", row.names = FALSE) %>%
    kable_styling(full_width=FALSE, font_size=14, position="left")
} else {
  cat("*Univariate OR table not available.*")
}
```

### Model Coefficients

**Logistic regression coefficients, standard errors, and z-values. Model includes final 4 predictors.**

```{r results='asis'}
husbandry_coeffs <- as.data.frame(summary_fit_husbandry$coefficients)
names(husbandry_coeffs) <- c("Estimate", "Std. Error", "z value", "P-value")
husbandry_coeffs$Predictor <- decode_html(rownames(husbandry_coeffs))
rownames(husbandry_coeffs) <- NULL
husbandry_coeffs <- husbandry_coeffs[, c("Predictor", "Estimate", "Std. Error", "z value", "P-value")]
kable(
  husbandry_coeffs[,c("Predictor", "Estimate", "Std. Error", "z value")], digits = 3, caption = "Coefficients (Husbandry Model)", align = "l", row.names = FALSE
) %>%
  kable_styling(full_width=FALSE, font_size=14, position="left")
```

### Odds Ratios and Confidence Intervals

```{r results='asis'}
or_ci_husbandry_df <- as.data.frame(or_ci_husbandry)
names(or_ci_husbandry_df)[1] <- "Odds Ratio"
or_ci_husbandry_df$Predictor <- decode_html(rownames(or_ci_husbandry_df))
rownames(or_ci_husbandry_df) <- NULL
or_ci_husbandry_df <- or_ci_husbandry_df[, c("Predictor", "Odds Ratio", "2.5 %", "97.5 %")]
kable(
  or_ci_husbandry_df, digits=3, caption="Odds Ratios & 95% CI (Husbandry Model)", align = "l", row.names = FALSE
) %>%
  kable_styling(full_width=FALSE, font_size=14, position="left")
```

### Diagnostics Overview

<div style="font-size:1.12em; background:#f8f8f8; padding:14px; border-radius:8px; max-width:650px; margin-left:0;">
  <ul style="margin-bottom:0;">
    <li>
      <b>Pseudo-R² (McFadden):</b>
      <span style="color:#004488; font-weight:600;">`r round(pseudo_r2_husbandry, 3)`</span>
      &mdash; &gt; 0.2 = reasonable; lower = weak fit.<br>
      <span style="font-size:93%; color:#555;">
        `r if (pseudo_r2_husbandry > 0.2) {"Model fit is reasonable."} else {"Model fit is weak."}`
      </span>
    </li>
    <li>
      <b>AUC (ROC Curve):</b>
      <span style="color:#004488; font-weight:600;">`r round(auc_value_husbandry, 3)`</span>
      &mdash; &gt; 0.7 = acceptable, &gt; 0.8 = good.<br>
      <span style="font-size:93%; color:#555;">
        `r if (auc_value_husbandry > 0.8) {"Discrimination is good."} else if (auc_value_husbandry > 0.7) {"Discrimination is acceptable."} else {"Discrimination is poor."}`
      </span>
    </li>
  </ul>
</div>

### Multicollinearity

**Variance Inflation Factor (VIF) for each predictor. VIF > 5 signals multicollinearity.**

```{r results='asis'}
vif_husbandry_df <- data.frame(Predictor = decode_html(names(vif_vals_husbandry)), VIF = as.vector(vif_vals_husbandry))
kable(
  vif_husbandry_df, digits=2, caption="Variance Inflation Factor (Husbandry Model)", align = "l", row.names = FALSE
) %>%
  kable_styling(full_width=FALSE, font_size=14, position="left")
```

### Top 5 Influential Observations

**Cook's Distance for the top 5 most influential points; high values may indicate outliers.**

```{r results='asis'}
cooks_husbandry_df <- data.frame(Observation = names(head(sort(cooksd_husbandry, decreasing = TRUE), 5)),
                                 Cook_Distance = head(sort(cooksd_husbandry, decreasing = TRUE), 5))
kable(
  cooks_husbandry_df, digits=3, caption="Top 5 Cook's Distance Values (Husbandry Model)", align = "l", row.names = FALSE
) %>%
  kable_styling(full_width=FALSE, font_size=14, position="left")
```

### Visualizations

```{r, fig.width=10, fig.height=4}
par(mfrow = c(1, 2))
plot(roc_obj_husbandry, main = "ROC Curve (Husbandry)", col = "blue", lwd = 2)
plot(cooksd_husbandry, type = "h", main = "Cook's Distance (Husbandry)", ylab = "Cook's D", xlab = "Observation")
abline(h = 4 / length(cooksd_husbandry), col = "red", lty = 2)
par(mfrow = c(1, 1))
```

### Calibration

**Calibration of model predictions was assessed by the Hosmer-Lemeshow test and a calibration plot.**

```{r calibration, results='asis', fig.width=6, fig.height=5}
if (exists("calib_hl_husbandry") && !is.null(calib_hl_husbandry)) {
  cat("<b>Hosmer-Lemeshow Test:</b> <br>")
  cat("Chi-squared =", signif(calib_hl_husbandry$statistic, 3), 
      ", df =", calib_hl_husbandry$parameter, 
      ", p-value =", signif(calib_hl_husbandry$p.value, 3), "<br>")
  cat("A p-value > 0.05 suggests good calibration.<br><br>")
} else {
  cat("*Hosmer-Lemeshow test not available.*<br>")
}

if (exists("calib_plot_husbandry") && !is.null(calib_plot_husbandry)) {
  calib_plot_husbandry()
} else if (exists("calib_rms_husbandry") && !is.null(calib_rms_husbandry)) {
  plot(calib_rms_husbandry, main = "Calibration Curve (Husbandry Model)")
} else {
  cat("*Calibration curve not available.*")
}
```

---

## Interpretation Notes

- **Exploratory Analysis:** No p-values or statistical significance are interpreted for selection, but all effect sizes and CIs are transparently reported.  
- **Biological plausibility exclusion:** See the 'exclusion_due_to_biological_plausibility' column in tables for rationale for exclusion.  
- **Model Fit:** Consider pseudo-R² and AUC for overall model quality.
- **Calibration:** Hosmer-Lemeshow and calibration curve should be checked for model reliability.
- **Multicollinearity:** High VIFs may require model simplification.
- **Influential Points:** Consider sensitivity analysis excluding high Cook’s D observations.

---

## Session Info

```{r session-info, echo=FALSE}
session_info_path <- file.path(output_dir, "SIV_logistic_regression_sessionInfo.txt")
if (file.exists(session_info_path)) {
  cat(readLines(session_info_path), sep="\n")
} else {
  sessionInfo()
}
```

---

<span style="font-size:16px;">
Prepared by <b>jonasalexandersteiner</b> on <b>`r format(Sys.Date())`</b>.
</span>Fl